Locally Conservative Discontinuous Petrov-Galerkin Finite Elements for Fluid Problems

### 2013-12-04 docs/Truman/LocalConservation/paper/paper.tex
Added punctuation after equations in LC paper

The discontinuous Petrov-Galerkin finite element method has been described as
least squares finite elements with a twist. The key difference is that least
squares methods seek to minimize the residual of the solution in the $L^2$
norm, while DPG seeks the minimization in a dual norm realized through the
inverse Riesz map. Exact mass conservation has been an issue that has plagued
least squares finite elements for a long time. Several approaches have been
used to try to address this. Bochev \etal\cite{Bochev2010} accomplish local
conservation by using a pointwise divergence free velocity space in the Stokes
formulation.  Chang and Nelson\cite{ChangNelson1997} developed the
\emph{restricted LSFEM}\cite{ChangNelson1997} by augmenting the least squares
equations with a Lagrange multiplier explicitly enforcing mass conservation
element-wise. Our conservative formulation of DPG takes a similar approach and
both methods share similar negative of transforming a minimization method to a
saddle point problem. In the interest of crediting Chang and Nelson's
restricted LSFEM, we call the following locally conservative DPG method the
restricted DPG method (RDPG).

\subsection{DPG is a Minimum Residual Method}
Roberts \etal presents a brief history and derivation of DPG with optimal test functions in
\cite{DPGStokes}. We follow his derivation of the standard DPG method as a
minimum residual method. Let $U$ be the trial Hilbert space and $V$ the test
Hilbert space for a well-posed variational problem $b(u,v)=l(v)$. In operator
form this is $Bu=l$, where $B:U\rightarrow V'$ and $\LRa{Bu,v}=b(u,v)$. We seek to minimize the
residual for the discrete space $U_h\subset U$:
\begin{equation}
u_h=\argmin_{u_h\in U_h}\frac{1}{2}\norm{Bu_h-l}^2_{V'}\,.
\label{minresidual}
\end{equation}
Recalling that the Riesz operator $R_V:V\rightarrow V'$ is an isometry defined
by
\[
\LRa{R_Vv,\delta v}=\LRp{v,\delta v}_V,\quad\forall\delta v\in V,
\]
we can use the Riesz inverse to minimize in the $V$-norm rather than its dual:
\begin{equation}
\frac{1}{2}\norm{Bu_h-l}^2_{V'}=\frac{1}{2}\norm{R_V^{-1}(Bu_h-l)}^2_V
=\frac{1}{2}\LRp{R_V^{-1}(Bu_h-l),R_V^{-1}(Bu_h-l)}_V\,.
\label{eq:rieszapplied}
\end{equation}
The first order optimality condition for \eqnref{rieszapplied} requires
the G\^ateaux derivative to be zero in all directions $\delta u \in
U_h$, i.e.,
\[
\left(R_V^{-1}(Bu_h-l),R_V^{-1}B\delta u\right)_V = 0, \quad \forall \delta u \in U.
\]
By definition of the Riesz operator, this is equivalent to
\begin{equation}
\LRa{Bu_h-l,R_V^{-1}B\delta u_h}=0\quad\forall\delta u_h\in U_h\,.
\label{eq:DPGbilinearform}
\end{equation}
Now, we can identify $v_{\delta u_h}\coloneqq R_V^{-1}B\delta u_h$ as the
optimal test function for trial function $\delta u_h$. Define $T:=R_V^{-1}B:U_h\rightarrow V$ as the trial-to-test operator. Now we can rewrite
\eqnref{DPGbilinearform} as
\begin{equation}
b(u_h,v_{\delta u_h})=l(v_{\delta u_h}).
\label{eq:DPGmethod}
\end{equation}
The DPG method then is to solve \eqnref{DPGmethod} with optimal test functions
$v_{\delta u_h}\in V$ that solve the auxiliary problem
\begin{equation}
\LRp{v_{\delta u_h},\delta v}_V=\LRa{R_Vv_{\delta u_h},\delta v}
=\LRa{B\delta u_h,\delta v}=b(\delta u_h,\delta v)\,,\quad\forall\delta v\in V.
\label{eq:optimaltestproblem}
\end{equation}
Using a continuous test basis would result in a global solve for every optimal
test function. Therefore DPG uses a discontinuous test basis which makes each
solve element-local and much more computationally tractable. Of course,
\eqnref{optimaltestproblem} still requires the inversion of the
infinite-dimensional Riesz map, but approximating $V$ by a finite
dimensional $V_h$ which is of a higher polynomial degree than $U_h$ (hence
``enriched space'') works well in practice.

@@

Let function $\bfpsi$ denote the $H(\text{div},\Omega)$ extension of flux $\hat{t}$
that realizes the minimum in the definition of the quotient (minimum energy
extension) norm.
The choice of norm for the Lagrange multipliers $\lambda_K$ is implied
by the quotient norm used for $H^{-1/2}(\Gamma_h)$ and continuity
bound for form $c(p,\bfw)$ representing the constraint:
\begin{equation}
\begin{array}{lll}
| c(\sum_K \lambda_K (1_K,{\bf 0}),(u,\bfsig,\hat{u},\hat{t})) |
& = | \sum_K \lambda_K \langle \hat{t}, 1_K \rangle_{\ptl K} | \\[8pt]
& = | \sum_K \lambda_K \langle v_n , 1_K \rangle_{\ptl K} | \\[8pt]
& = | \sum_K \lambda_K \int_K \text{div} \bfpsi \: 1_K  | \\[8pt]
& \leq \sum_K  \lambda_K || \text{div} \bfpsi ||_{L^2(K)} \mu(K)^{1/2} \\[8pt]
& \leq (\sum_K \mu(K) \lambda_K^2)^{1/2} \: (\sum_K || \text{div} \bfpsi ||_{L^2(K)}^2 )^{1/2} \\[8pt]
& \leq \underbrace{\left(\sum_K \mu(K) \lambda_K^2\right)^{1/2}}_{=: || \bflambda ||} ||
\hat{t} ||_{H^{-1/2}(\Gamma_h)}
\end{array}
\end{equation}
where $\mu(K)$ stands for the area (measure) of element $K$.

@@

\paragraph{Inf Sup Condition} relating spaces $\bfU$ and $Q$ reads as follows,
\begin{equation}
   \sup_{\bfw \in \bfU} \frac{| c(p,\bfw) |}{|| \bfw ||_{\bfU}} \geq \beta ||
   p ||_Q
\end{equation}
Let
\begin{equation}
R\, : \, L^2(\Omega) \ni q \to \bfpsi \in \bfH(\text{div},\Omega) \cap \bfH^1(\Omega)
=\bfH^1(\Omega)
\end{equation}
be
the continuous right inverse of the divergence operator constructed by
Costabel and McIntosh in \cite{CostabelMcIntosh}.
Let $\bfpsi_h$ denote the classical, lowest order Raviart-Thomas (RT) interpolant of
function
\begin{equation}
\bfpsi = R (\sum_K \lambda_K 1_K) \: .
\end{equation}
Note that $\text{div} \bfpsi_h = \text{div} \bfpsi = \lambda_K$ in element $K$.

Classical $h$-interpolation interpolation error estimate for the lowest error
Raviart-Thomas elements and continuity of operator $R$ imply the stability estimate:
\begin{equation}
\begin{array}{lll}
|| \bfpsi_h || & \leq || \bfpsi_h - \bfpsi || + || \bfpsi ||\\[8pt]
&\leq C h || \bfpsi ||_{H^1} +  || \bfpsi || \\[8pt]
& \leq C || \text{div} \bfpsi || = C (\sum_K \mu(K) \lambda_K^2)^{1/2}
\end{array}
\label{eq:stab}
\end{equation}
Above, $C$ is a generic, mesh independent constant incorporating constant from
the interpolation error estimate and continuity constant of $R$.
Let $\hat{t}$ be the trace of $\bfpsi_h$. We have then,
\begin{equation}
\sup_{\hat{t} \in H^{-1/2}(\Gamma_h)} \frac{|  \sum_K \lambda_K \langle
\hat{t},1_K \rangle_{\ptl K} |}{|| \hat{t} ||_{H^{-1/2}(\Gamma_h)}}
\geq \frac{| \sum_K \lambda_K \int_K \text{div} \bfpsi_h \: 1_K  |}
{|| \bfpsi_h ||_{H(\text{div},\Omega)}}
\geq \frac{1}{C} (\sum_K \mu(K) \lambda_K^2)^{1/2}
\end{equation}
where $C$ is the constant from stability estimate~(\ref{eq:stab}).

@@

\paragraph{Inf Sup in Kernel Condition} is satisfied automatically due to the use of optimal
test functions. First of all, we characterize the ``kernel'' space:
\begin{equation}
\begin{array}{rl}
\bfU_0  := & \{ \bfw \in \bfU \, : \, c(q,\bfw) = 0 \quad \forall q \in Q\} \\[8pt]
 = &\{ (u,\bfsig,\hat{u},\hat{t}) \, : \, \langle \hat{t},1_K \rangle = 0
 \quad \forall K \}
\end{array}
\end{equation}
In other words, the kernel space contains only the equilibrated fluxes.
With $\bfu \in \bfU_0$, we have then:
\begin{equation}
   \sup_{\bfw \in \bfU_0} \frac{| a(\bfu,\bfw) |}{|| \bfw ||_{\bfU} }
   \geq \frac{| b(\bfu,T \bfu) |}{|| \bfu ||}
   = \frac{| b(\bfu,T \bfu) |}{|| T\bfu ||}\frac{||T\bfu||}{||\bfu||}
   = \sup_{(v,\bftau)}\frac{| b((u,\bfsig,\hat{u},\hat{t}), (v,\bftau)) |}{|| (v,\bftau) ||}
   \frac{||T\bfu||}{||\bfu||}
   \geq \gamma^2 || (u,\bfsig,\hat{u},\hat{t}) ||
\end{equation}
where $\gamma$ is the stability constant for the standard continuous DPG formulation.
The first inequality follows as we plug in the definition for $a$ and pick
$\bfw=\bfu$. The second equality is trivial, while the next one follows by definition of the optimal test
functions given through the trial-to-test operator $T$. The finally inequality
springs from the fact that
$\sup_{\bfv}\frac{|b(\bfu,\bfv)|}{||\bfv||}\geq\gamma||\bfu||$ and
$||T\bfu||_V=||R_V^{-1}B\bfu||_V=||B\bfu||_{V'}\geq\gamma||u||$.

@@

Let us start with the inf-sup condition,
\begin{equation}
\sup_{\hat{t}} \frac{| \sum_K \lambda_K \langle \hat{t},1_K \rangle |}{|| \hat{t} ||_E}
\geq  \beta \LRp{\sum_K \mu(K) \lambda_K^2}^{1/2}\,.
\end{equation}
As in the previous analysis, we select for $\hat{t}$ the trace of Raviart-Thomas
interpolant $\bfpsi_h$ of $\bfpsi = R (\sum_K \lambda_K 1_K)$ where $R$ is the right-inverse
of the divergence operator constructed by Costabel and McIntosh. The only change compared
with the previous analysis, is the evaluation of norm of $\hat{t}_h$. For this, we need
to solve the local problems:
\begin{align}
((v,\bftau), (\delta v, \delta \bftau)_V &= \langle \hat{t}, \delta v \rangle_{\ptl K}
= \int_K \text{div} \bfpsi_h \, \delta v = \int_K \text{div} \bfpsi \, \delta v
\nonumber\\
&= \int_K \lambda_K \delta v
= \lambda_K (1_K,\delta v)_K  \quad \forall \delta v \in
H^1(K)\,\forall\delta\bftau\in\bfH(\text{div},K)
\label{eq:local}
\end{align}
We need then an upper bound of the energy norm of $(v_h,\bftau_h)$:
$$
\LRp{\sum_K || (v,\bftau) ||_V^2}^{1/2}\,.
$$
Substituting $(v,\bftau)$ for $(\delta v,\delta\bftau)$ in~(\ref{eq:local}), we get,
\begin{equation}
|| (v,\bftau) ||_V^2 = \lambda_K (1_K, v_K)
\end{equation}
If we have a robust stability estimate:
\begin{equation}
(1_K, v_K) \leq C \mu(K)^{1/2} || (v,\bftau) ||_K
\label{eq:robustEst1}
\end{equation}
(i.e. constant $C$ is independent of $\epsilon$) then
\begin{equation}
||  (v,\bftau) ||_V \leq C \mu(K)^{1/2} | \lambda_K |
\end{equation}
and, eventually, as needed,
\begin{equation}
\sum_K || (v,\bftau) ||_V^2  \leq C^2 \sum_K \mu(K) \lambda_K^2
\end{equation}
which leads to the robust estimate of inf-sup constant $\beta$. For example, it is sufficient if
\begin{equation}
|| v ||_{L^2(K)} \leq || (v,\bftau) ||_V \, .
\label{eq:robust_est1}
\end{equation}
Notice that the stability analysis with the energy norm was, in a sense, easier than
with the quotient norm. Only the divergence of the interpolant $\bfpsi_h$ enters~(\ref{eq:local})
and it coincides with the divergence of $\bfpsi$.

We arrive at a similar situation in the continuity estimate of
$$
\sum_K \lambda_K \langle \hat{t}, 1_K \rangle\,.
$$
Testing with $(1_K,{\bf 0})$ in the local problem~(\ref{eq:local_problem}), we obtain,
\begin{equation}
((v,\bftau),(1_K,{\bf 0}))_V = \langle \hat{t}, 1_K \rangle_{\ptl K}
\label{eq:local2}
\end{equation}
If we have a robust estimate,
\begin{equation}
| ((v,\bftau),(1_K,{\bf 0}))_V | \leq C \mu(K)^{1/2} \, || (v,\bftau) ||_V
\label{eq:robust_est2}
\end{equation}
then
\begin{equation}
| \sum_K \lambda_K \langle \hat{t}, 1_K \rangle | \leq C (\sum_K \mu(K) \lambda_K^2)^{1/2}
\, (\sum_K || (v,\bftau) ||_V^2)^{1/2}
= C (\sum_K \mu(K) \lambda_K^2)^{1/2}  || \hat{t} ||_E
\end{equation}
as needed.

@@

While oscillations of this sort in a solution near a singular point may be
acceptable in certain simulations, it is a big problem for the methods in
compressible flow simulations -- physical constraints require several solution
variables to remain positive throughout simulation.\footnote{Apart from
returning a non-physical solution, the violation of positivity constraints
typically results in non-convergence of nonlinear solvers.}  We propose a
modification of the robust test norm that appears to remedy this issue, which
we refer to as the coupled robust test norm:
\begin{align}
\norm{\LRp{v,\btau}}_{V,K}^2 &\coloneqq
\min\left\{\frac{1}{\epsilon},\frac{1}{\mu(K)}\right\}\norm{\btau}_K^2,
+ \norm{\div \btau - \bbeta\cdot\Grad v}_K^2 \\
&+\norm{\bbeta\cdot \Grad v}_K^2
+ \epsilon\norm{\Grad v}_K^2
+ \norm{v}^2_K
\end{align}
where $||\cdot||_K$ signifies the $L^2$ norm over element $K$.\footnote{
We note that we have dropped the
mesh-dependent scaling on $\norm{v}_{\L}$ from the robust norm; this is related
to recent insights into the nature of DPG test spaces as discussed in
\cite{Chan2013}.
}
We note that, under the
theory developed in \cite{ChanHeuerThanhDemkowicz2012}
the above test norm is
trivially provably robust using the same ideas.\footnote{This is due to the
fact that $\norm{\div \btau - \bbeta\cdot\Grad v}_{\L}^2$ is robustly bounded by
$\norm{\div \btau}_{\L}$ and $\norm{\bbeta\cdot\Grad v}_{\L}$.  Alternatively, we
can note that $\norm{\div \btau - \bbeta\cdot\Grad v}_{\L}^2 = \norm{g}_{\L}^2$,
where $g$ is a load of the adjoint problem related to robustness described in
\cite{ChanHeuerThanhDemkowicz2012}.
}

@@

With a locally conservative formulation, we can pass in local problem
\eqref{eq:localSolve} with $\alpha\rightarrow0$. The fact that the test
functions will be determined then up to a constant does not matter, for
$\delta\hat t\in\hat F_h^e$, equation
\eqref{eq:conservativeConfusionSystem}$_1$ is orthogonal to constants.
Mathematically, we are dealing with equivalence classes of functions, but in
order to obtain a single function that we can deal with numerically, we
replace the alpha term with a zero mean scaling condition to obtain the new
test norm,
\begin{align}
\label{eq:localSolveMod}
\min&\left\{\frac{1}{\epsilon},\frac{1}{\mu(K)}\right\}(\taudeltau,\delta\btau)_K
+(\nabla\cdot\taudeltau-\bbeta\cdot\nabla
v,\nabla\cdot\delta\btau-\bbeta\cdot\nabla v)_K\\
&+(\bbeta\cdot\nabla\vdeltau,\bbeta\cdot\nabla \delta v)_K\nonumber
+\epsilon(\nabla\vdeltau,\nabla \delta v)_K
+\frac{1}{\mu(K)}\int_K\vdeltau\int_K\delta v
\end{align}
where the $\frac{1}{\mu(K)}$ coefficient is an arbitrary scaling condition
that doesn't make a difference mathematically, but can affect the condition
number of the actual solve. In practice, we use $\frac{1}{\mu(K)^2}$
since $\int_K\vdeltau$ and $\int_K\delta v$ both scale like
$\mu(K)$, but $\frac{1}{\mu(K)}$ is more convenient more the analysis in the
next section.
It is convenient to be able to take $\alpha\rightarrow0$ as we will see in
some later numerical experiments.

>>>>

The discontinuous Petrov-Galerkin finite element method has been described as
least squares finite elements with a twist. The key difference is that least
squares methods seek to minimize the residual of the solution in the $L^2$
norm, while DPG seeks the minimization in a dual norm realized through the
inverse Riesz map. Exact mass conservation has been an issue that has plagued
least squares finite elements for a long time. Several approaches have been
used to try to address this. Bochev \etal\cite{Bochev2010} accomplish local
conservation by using a point-wise divergence free velocity space in the Stokes
formulation.  Chang and Nelson\cite{ChangNelson1997} developed the
\emph{restricted LSFEM}\cite{ChangNelson1997} by augmenting the least squares
equations with a Lagrange multiplier explicitly enforcing mass conservation
element-wise. Our conservative formulation of DPG takes a similar approach and
both methods share similar negative of transforming a minimization method to a
saddle point problem. In the interest of crediting Chang and Nelson's
restricted LSFEM, we call the following locally conservative DPG method the
restricted DPG method (RDPG).

\subsection{DPG is a Minimum Residual Method}
Roberts \etal presents a brief history and derivation of DPG with optimal test functions in
\cite{DPGStokes}. We follow his derivation of the standard DPG method as a
minimum residual method. Let $U$ be the trial Hilbert space and $V$ the test
Hilbert space for a well-posed variational problem $b(u,v)=l(v)$. In operator
form this is $Bu=l$, where $B:U\rightarrow V'$ and $\LRa{Bu,v}=b(u,v)$. We seek to minimize the
residual for the discrete space $U_h\subset U$:
\begin{equation}
u_h=\argmin_{u_h\in U_h}\frac{1}{2}\norm{Bu_h-l}^2_{V'}\,.
\label{minresidual}
\end{equation}
Recalling that the Riesz operator $R_V:V\rightarrow V'$ is an isometry defined
by
\[
\LRa{R_Vv,\delta v}=\LRp{v,\delta v}_V,\quad\forall\delta v\in V,
\]
we can use the Riesz inverse to minimize in the $V$-norm rather than its dual:
\begin{equation}
\frac{1}{2}\norm{Bu_h-l}^2_{V'}=\frac{1}{2}\norm{R_V^{-1}(Bu_h-l)}^2_V
=\frac{1}{2}\LRp{R_V^{-1}(Bu_h-l),R_V^{-1}(Bu_h-l)}_V\,.
\label{eq:rieszapplied}
\end{equation}
The first order optimality condition for \eqnref{rieszapplied} requires
the G\^ateaux derivative to be zero in all directions $\delta u \in
U_h$, i.e.,
\[
\left(R_V^{-1}(Bu_h-l),R_V^{-1}B\delta u\right)_V = 0, \quad \forall \delta u \in U.
\]
By definition of the Riesz operator, this is equivalent to
\begin{equation}
\LRa{Bu_h-l,R_V^{-1}B\delta u_h}=0\quad\forall\delta u_h\in U_h\,.
\label{eq:DPGbilinearform}
\end{equation}
Now, we can identify $v_{\delta u_h}\coloneqq R_V^{-1}B\delta u_h$ as the
optimal test function for trial function $\delta u_h$. Define $T:=R_V^{-1}B:U_h\rightarrow V$ as the trial-to-test operator. Now we can rewrite
\eqnref{DPGbilinearform} as
\begin{equation}
b(u_h,v_{\delta u_h})=l(v_{\delta u_h}).
\label{eq:DPGmethod}
\end{equation}
The DPG method then is to solve \eqnref{DPGmethod} with optimal test functions
$v_{\delta u_h}\in V$ that solve the auxiliary problem
\begin{equation}
\LRp{v_{\delta u_h},\delta v}_V=\LRa{R_Vv_{\delta u_h},\delta v}
=\LRa{B\delta u_h,\delta v}=b(\delta u_h,\delta v)\quad\forall\delta v\in V.
\label{eq:optimaltestproblem}
\end{equation}
Using a continuous test basis would result in a global solve for every optimal
test function. Therefore DPG uses a discontinuous test basis which makes each
solve element-local and much more computationally tractable. Of course,
\eqnref{optimaltestproblem} still requires the inversion of the
infinite-dimensional Riesz map, but approximating $V$ by a finite
dimensional $V_h$ which is of a higher polynomial degree than $U_h$ (hence
``enriched space'') works well in practice.

@@

Let function $\bfpsi$ denote the $H(\text{div},\Omega)$ extension of flux $\hat{t}$
that realizes the minimum in the definition of the quotient (minimum energy
extension) norm.
The choice of norm for the Lagrange multipliers $\lambda_K$ is implied
by the quotient norm used for $H^{-1/2}(\Gamma_h)$ and continuity
bound for form $c(p,\bfw)$ representing the constraint:
\begin{equation}
\begin{array}{lll}
| c(\sum_K \lambda_K (1_K,{\bf 0}),(u,\bfsig,\hat{u},\hat{t})) |
& = | \sum_K \lambda_K \langle \hat{t}, 1_K \rangle_{\ptl K} | \\[8pt]
& = | \sum_K \lambda_K \langle v_n , 1_K \rangle_{\ptl K} | \\[8pt]
& = | \sum_K \lambda_K \int_K \text{div} \bfpsi \: 1_K  | \\[8pt]
& \leq \sum_K  \lambda_K || \text{div} \bfpsi ||_{L^2(K)} \mu(K)^{1/2} \\[8pt]
& \leq (\sum_K \mu(K) \lambda_K^2)^{1/2} \: (\sum_K || \text{div} \bfpsi ||_{L^2(K)}^2 )^{1/2} \\[8pt]
& \leq \underbrace{\left(\sum_K \mu(K) \lambda_K^2\right)^{1/2}}_{=: || \bflambda ||} ||
\hat{t} ||_{H^{-1/2}(\Gamma_h)}\,,
\end{array}
\end{equation}
where $\mu(K)$ stands for the area (measure) of element $K$.

@@

\paragraph{Inf Sup Condition} relating spaces $\bfU$ and $Q$ reads as follows:
\begin{equation}
   \sup_{\bfw \in \bfU} \frac{| c(p,\bfw) |}{|| \bfw ||_{\bfU}} \geq \beta ||
   p ||_Q\,.
\end{equation}
Let
\begin{equation}
R\, : \, L^2(\Omega) \ni q \to \bfpsi \in \bfH(\text{div},\Omega) \cap \bfH^1(\Omega)
=\bfH^1(\Omega)
\end{equation}
be
the continuous right inverse of the divergence operator constructed by
Costabel and McIntosh in \cite{CostabelMcIntosh}.
Let $\bfpsi_h$ denote the classical, lowest order Raviart-Thomas (RT) interpolant of
function
\begin{equation}
\bfpsi = R (\sum_K \lambda_K 1_K) \: .
\end{equation}
Note that $\text{div} \bfpsi_h = \text{div} \bfpsi = \lambda_K$ in element $K$.

Classical $h$-interpolation interpolation error estimate for the lowest error
Raviart-Thomas elements and continuity of operator $R$ imply the stability estimate:
\begin{equation}
\begin{array}{lll}
|| \bfpsi_h || & \leq || \bfpsi_h - \bfpsi || + || \bfpsi ||\\[8pt]
&\leq C h || \bfpsi ||_{H^1} +  || \bfpsi || \\[8pt]
& \leq C || \text{div} \bfpsi || = C (\sum_K \mu(K) \lambda_K^2)^{1/2}\,.
\end{array}
\label{eq:stab}
\end{equation}
Above, $C$ is a generic, mesh independent constant incorporating constant from
the interpolation error estimate and continuity constant of $R$.
Let $\hat{t}$ be the trace of $\bfpsi_h$. We have then,
\begin{equation}
\sup_{\hat{t} \in H^{-1/2}(\Gamma_h)} \frac{|  \sum_K \lambda_K \langle
\hat{t},1_K \rangle_{\ptl K} |}{|| \hat{t} ||_{H^{-1/2}(\Gamma_h)}}
\geq \frac{| \sum_K \lambda_K \int_K \text{div} \bfpsi_h \: 1_K  |}
{|| \bfpsi_h ||_{H(\text{div},\Omega)}}
\geq \frac{1}{C} (\sum_K \mu(K) \lambda_K^2)^{1/2}\,,
\end{equation}
where $C$ is the constant from stability estimate~(\ref{eq:stab}).

@@

\paragraph{Inf Sup in Kernel Condition} is satisfied automatically due to the use of optimal
test functions. First of all, we characterize the ``kernel'' space:
\begin{equation}
\begin{array}{rl}
\bfU_0  := & \{ \bfw \in \bfU \, : \, c(q,\bfw) = 0 \quad \forall q \in Q\} \\[8pt]
 = &\{ (u,\bfsig,\hat{u},\hat{t}) \, : \, \langle \hat{t},1_K \rangle = 0
 \quad \forall K \}\,.
\end{array}
\end{equation}
In other words, the kernel space contains only the equilibrated fluxes.
With $\bfu \in \bfU_0$, we have then:
\begin{equation}
   \sup_{\bfw \in \bfU_0} \frac{| a(\bfu,\bfw) |}{|| \bfw ||_{\bfU} }
   \geq \frac{| b(\bfu,T \bfu) |}{|| \bfu ||}
   = \frac{| b(\bfu,T \bfu) |}{|| T\bfu ||}\frac{||T\bfu||}{||\bfu||}
   = \sup_{(v,\bftau)}\frac{| b((u,\bfsig,\hat{u},\hat{t}), (v,\bftau)) |}{|| (v,\bftau) ||}
   \frac{||T\bfu||}{||\bfu||}
   \geq \gamma^2 || (u,\bfsig,\hat{u},\hat{t}) ||\,,
\end{equation}
where $\gamma$ is the stability constant for the standard continuous DPG formulation.
The first inequality follows as we plug in the definition for $a$ and pick
$\bfw=\bfu$. The second equality is trivial, while the next one follows by definition of the optimal test
functions given through the trial-to-test operator $T$. The finally inequality
springs from the fact that
$\sup_{\bfv}\frac{|b(\bfu,\bfv)|}{||\bfv||}\geq\gamma||\bfu||$ and
$||T\bfu||_V=||R_V^{-1}B\bfu||_V=||B\bfu||_{V'}\geq\gamma||u||$.

@@

Let us start with the inf-sup condition,
\begin{equation}
\sup_{\hat{t}} \frac{| \sum_K \lambda_K \langle \hat{t},1_K \rangle |}{|| \hat{t} ||_E}
\geq  \beta \LRp{\sum_K \mu(K) \lambda_K^2}^{1/2}\,.
\end{equation}
As in the previous analysis, we select for $\hat{t}$ the trace of Raviart-Thomas
interpolant $\bfpsi_h$ of $\bfpsi = R (\sum_K \lambda_K 1_K)$ where $R$ is the right-inverse
of the divergence operator constructed by Costabel and McIntosh. The only change compared
with the previous analysis, is the evaluation of norm of $\hat{t}_h$. For this, we need
to solve the local problems:
\begin{align}
((v,\bftau), (\delta v, \delta \bftau)_V &= \langle \hat{t}, \delta v \rangle_{\ptl K}
= \int_K \text{div} \bfpsi_h \, \delta v = \int_K \text{div} \bfpsi \, \delta v
\nonumber\\
&= \int_K \lambda_K \delta v
= \lambda_K (1_K,\delta v)_K  \quad \forall \delta v \in
H^1(K)\,\forall\delta\bftau\in\bfH(\text{div},K)\,.
\label{eq:local}
\end{align}
We need then an upper bound of the energy norm of $(v_h,\bftau_h)$:
$$
\LRp{\sum_K || (v,\bftau) ||_V^2}^{1/2}\,.
$$
Substituting $(v,\bftau)$ for $(\delta v,\delta\bftau)$ in~(\ref{eq:local}), we get,
\begin{equation}
|| (v,\bftau) ||_V^2 = \lambda_K (1_K, v_K)\,.
\end{equation}
If we have a robust stability estimate:
\begin{equation}
(1_K, v_K) \leq C \mu(K)^{1/2} || (v,\bftau) ||_K
\label{eq:robustEst1}
\end{equation}
(i.e. constant $C$ is independent of $\epsilon$), then
\begin{equation}
||  (v,\bftau) ||_V \leq C \mu(K)^{1/2} | \lambda_K |
\end{equation}
and, eventually as needed,
\begin{equation}
\sum_K || (v,\bftau) ||_V^2  \leq C^2 \sum_K \mu(K) \lambda_K^2\,,
\end{equation}
which leads to the robust estimate of inf-sup constant $\beta$. For example, it is sufficient if
\begin{equation}
|| v ||_{L^2(K)} \leq || (v,\bftau) ||_V \, .
\label{eq:robust_est1}
\end{equation}
Notice that the stability analysis with the energy norm was, in a sense, easier than
with the quotient norm. Only the divergence of the interpolant $\bfpsi_h$ enters~(\ref{eq:local})
and it coincides with the divergence of $\bfpsi$.

We arrive at a similar situation in the continuity estimate of
$$
\sum_K \lambda_K \langle \hat{t}, 1_K \rangle\,.
$$
Testing with $(1_K,{\bf 0})$ in the local problem~(\ref{eq:local_problem}), we obtain,
\begin{equation}
((v,\bftau),(1_K,{\bf 0}))_V = \langle \hat{t}, 1_K \rangle_{\ptl K}\,.
\label{eq:local2}
\end{equation}
If we have a robust estimate,
\begin{equation}
| ((v,\bftau),(1_K,{\bf 0}))_V | \leq C \mu(K)^{1/2} \, || (v,\bftau) ||_V\,,
\label{eq:robust_est2}
\end{equation}
then
\begin{equation}
| \sum_K \lambda_K \langle \hat{t}, 1_K \rangle | \leq C (\sum_K \mu(K) \lambda_K^2)^{1/2}
\, (\sum_K || (v,\bftau) ||_V^2)^{1/2}
= C (\sum_K \mu(K) \lambda_K^2)^{1/2}  || \hat{t} ||_E\,,
\end{equation}
as needed.

@@

While oscillations of this sort in a solution near a singular point may be
acceptable in certain simulations, it is a big problem for the methods in
compressible flow simulations -- physical constraints require several solution
variables to remain positive throughout simulation.\footnote{Apart from
returning a non-physical solution, the violation of positivity constraints
typically results in non-convergence of nonlinear solvers.}  We propose a
modification of the robust test norm that appears to remedy this issue, which
we refer to as the coupled robust test norm:
\begin{align}
\norm{\LRp{v,\btau}}_{V,K}^2 &\coloneqq
\min\left\{\frac{1}{\epsilon},\frac{1}{\mu(K)}\right\}\norm{\btau}_K^2,
+ \norm{\div \btau - \bbeta\cdot\Grad v}_K^2 \\
&+\norm{\bbeta\cdot \Grad v}_K^2
+ \epsilon\norm{\Grad v}_K^2
+ \norm{v}^2_K\,,
\end{align}
where $||\cdot||_K$ signifies the $L^2$ norm over element $K$.\footnote{
We note that we have dropped the
mesh-dependent scaling on $\norm{v}_{\L}$ from the robust norm; this is related
to recent insights into the nature of DPG test spaces as discussed in
\cite{Chan2013}.
}
We note that, under the
theory developed in \cite{ChanHeuerThanhDemkowicz2012}
the above test norm is
trivially provably robust using the same ideas.\footnote{This is due to the
fact that $\norm{\div \btau - \bbeta\cdot\Grad v}_{\L}^2$ is robustly bounded by
$\norm{\div \btau}_{\L}$ and $\norm{\bbeta\cdot\Grad v}_{\L}$.  Alternatively, we
can note that $\norm{\div \btau - \bbeta\cdot\Grad v}_{\L}^2 = \norm{g}_{\L}^2$,
where $g$ is a load of the adjoint problem related to robustness described in
\cite{ChanHeuerThanhDemkowicz2012}.
}

@@

With a locally conservative formulation, we can pass in local problem
\eqref{eq:localSolve} with $\alpha\rightarrow0$. The fact that the test
functions will be determined then up to a constant does not matter, for
$\delta\hat t\in\hat F_h^e$, equation
\eqref{eq:conservativeConfusionSystem}$_1$ is orthogonal to constants.
Mathematically, we are dealing with equivalence classes of functions, but in
order to obtain a single function that we can deal with numerically, we
replace the alpha term with a zero mean scaling condition to obtain the new
test norm,
\begin{align}
\label{eq:localSolveMod}
\min&\left\{\frac{1}{\epsilon},\frac{1}{\mu(K)}\right\}(\taudeltau,\delta\btau)_K
+(\nabla\cdot\taudeltau-\bbeta\cdot\nabla
v,\nabla\cdot\delta\btau-\bbeta\cdot\nabla v)_K\\
&+(\bbeta\cdot\nabla\vdeltau,\bbeta\cdot\nabla \delta v)_K\nonumber
+\epsilon(\nabla\vdeltau,\nabla \delta v)_K
+\frac{1}{\mu(K)}\int_K\vdeltau\int_K\delta v\,,
\end{align}
where the $\frac{1}{\mu(K)}$ coefficient is an arbitrary scaling condition
that doesn't make a difference mathematically, but can affect the condition
number of the actual solve. In practice, we use $\frac{1}{\mu(K)^2}$
since $\int_K\vdeltau$ and $\int_K\delta v$ both scale like
$\mu(K)$, but $\frac{1}{\mu(K)}$ is more convenient more the analysis in the
next section.
It is convenient to be able to take $\alpha\rightarrow0$ as we will see in
some later numerical experiments.
